### 2021-03-18 papers/SYRCoSE/encode_evstr.tex
feat: add part of small-step semantics section

\subsection{How we encode Event Structures}
% Как мы видели ранее в Step-by-step example наша стуркура событий содержит такие функции как frf fpred. Сами события это элементы некоторого множества E (множества итендификаторов), которое должно удволетворять следущим свойствам:
\begin{definition}[identifier set]
  $E$ is an identifier set if:
  \begin{itemize}
    \item There is some $ident_0 \in E$ ~--- first identifier
    \item $E$ has well-founded partitial order $\leqslant_E$
    \item There is some function, that takes some identifier $e$ and returns next fresh identifier. Formally: $fresh : E \to E$, satisfies $$\forall e \in E, \ e <_E fresh(e)$$
  \end{itemize}
\end{definition}
% Так же у нас есть функция которая каждому итендификатору сопоставляет его метку
  $$\ffun{lab} : E \rightharpoonup \lb $$
% где label -- множество всех меток (таких как W(x, 5), R(y, 6), ThreadStart, ThreadEnd etc)
% Мы будем говорить что е это "read"-event c локацией х и значением i если lab(e) = R(x,i), аналогично "write"-event
% В итоге формально мы строим четверку
\begin{definition}[Execution Event Structure]
  $$exec\_event\_structure \ \triangleq \ \langle E, \ \ffun{lab}, \ \fpred, \ \frf \rangle$$
  where
  \begin{itemize}
    \item $E$ ~--- identifiers set
    \item $\fpred$ ~--- function that returns a predsessor (w.r.t. the program order) of event
    \item $\frf$ ~--- function that takes "read"-event $e$ and returns a "write"-event from witch $e$ reads
    \item $\ffun{lab}$ ~--- function that returns event's label
  \end{itemize}
  Also we require
  \begin{itemize}
    \item $\forall e \in E, \ \fpred(e) \leqslant_E e$ and $\frf(e) \leqslant_E e$
    \item $\forall e_1 \ e_2 \in E, \ e_2 = \frf(e_1) \ \Leftrightarrow \ e_1$ is a "read"-event and $e_2$ is a "write"-event with the same location and value
  \end{itemize}
\end{definition}

>>>>

\subsection{How we encode Event Structures}
% Как мы видели ранее в Step-by-step example наша стуркура событий содержит такие функции как frf fpred. Сами события это элементы некоторого множества E (множества итендификаторов), которое должно удволетворять следущим свойствам:
\begin{definition}[identifier set]
  $\mathcal{E}$ is an identifier set if:
  \begin{itemize}
    \item There is some $ident_0 \in \mathcal{E}$ ~--- first identifier
    \item $\mathcal{E}$ has well-founded partitial order $\leqslant_\mathcal{E}$
    \item There is some function, that takes some finite identifies set $E \subset \mathcal{E}$ and returns a fresh identifier. Formally: $fresh : \mathcal{P}_{fin}(\mathcal{E}) \to \mathcal{E}$, satisfies $$\forall e \in E, \ e <_\mathcal{E} fresh(E)$$
  \end{itemize}
\end{definition}
% Пусть E это множества всех событий программы тогда у нас есть
  $$\ffun{lab} : E \rightharpoonup \lb $$
% где label -- множество всех меток (таких как W(x, 5), R(y, 6), ThreadStart, ThreadEnd etc)
% Мы будем говорить что е это "read"-event c локацией х и значением i если lab(e) = R(x,i), аналогично "write"-event
% В итоге формально мы строим четверку
\begin{definition}[Execution Event Structure]
  $$exec\_event\_structure \ \triangleq \ \langle E, \ \ffun{lab}, \ \fpred, \ \frf \rangle$$
  where
  \begin{itemize}
    \item $E$ ~--- finite subset of some identifiers set $\mathcal{E}$
    \item $\fpred$ ~--- function that returns a predsessor (w.r.t. the program order) of event
    \item $\frf$ ~--- function that takes "read"-event $e$ and returns a "write"-event from witch $e$ reads
    \item $\ffun{lab} : E \to \lb$ ~--- function that returns event's label
  \end{itemize}
  Also we require
  \begin{itemize}
    \item $\forall e \in E, \ \fpred(e) \leqslant_E e$ and $\frf(e) \leqslant_E e$
    \item $\forall e_1 \ e_2 \in E, \ e_2 = \frf(e_1) \ \Leftrightarrow \ e_1$ is a "read"-event and $e_2$ is a "write"-event with the same location and value
  \end{itemize}
\end{definition}

### 2021-03-18 papers/SYRCoSE/regmachine.tex
feat: add part of small-step semantics section

\subsection{Register Machine and thread semantics}
Now we are going to define language of simple register machine with shared memory.
\newcommand{\ip}{\ffun{ip}}
\newcommand{\regmap}{\ffun{regmap}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\[
\begin{array}{rcll}
  \texttt{P} &::=& \texttt{ST}^* \\
  \texttt{ST} & ::= & \texttt{I}^* \\
  \texttt{I} & ::= & \texttt{Reg := Val} \\
                 & | & \texttt{Loc := Val} \\
                 & | & \texttt{Reg := Loc} \\
                 & | & \texttt{CJmp Reg $\mathbb{N}$} \\
  \texttt{Reg}            & ::= & r \, | \, r' \, | ... \\ 
  \texttt{Loc}            & ::= & x \, | \,y \,| ... \\ 
\end{array}
\] where
\begin{itemize}
  \item \texttt{P} -- program is a sequence of single thread progams
  \item \texttt{ST} -- single thread program is a sequence of instructions
  \item \texttt{I} -- instruction can be either write to some register of some value or location value, either write to some location or conditional jump
\end{itemize}
Here \texttt{Reg} are registers (denoted as $r,r',...$) ~--- variables that should be different for different single thread progams. In contrast \texttt{Loc} are local variables (denoted as $x,y,...$) ~--- this is our shared (between different threads) memory. Also \texttt{Val} is the type of values (denoted as $i,j,...$) in our program, for simplicity let it be $\mathbb{N}$ or $\mathbb{Z}$. \\
We are interested at looking after shared memory, so in our event structures we have events that represent actions only at local variables. \\
Next we define semantics of single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define thread state:
\begin{definition}[Thread State]
  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \ffun{regmap} \rangle,$$ where
  \begin{itemize}
    \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
    \item $\ffun{regmap}$ ~--- map from registers to values
  \end{itemize}
\end{definition}
If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$. \\
Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where
\begin{itemize}
  \item $st$ ~--- thread state
  \item $p$ ~--- our single-thead programm
  \item $l$ ~--- label that denotes action on shared memory (can be with hole)
  \item $cont\_st$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
\end{itemize}
If we have some thread state $st$, we can identify $\lambda\, x,\, st$ (ie. constant continuation) with $st$ itself. Also by $f[r \leftarrow i]$ we will denote function that sends $r$ to $i$ and $r'$ to $f(r')$, for all $r' \neq r$

>>>>

\subsection{Register Machine and thread semantics}
Now we are going to define language of simple register machine with shared memory.
\[
\begin{array}{rcll}
  \texttt{P} &::=& \texttt{ST}^* \\
  \texttt{ST} & ::= & \texttt{I}^* \\
  \texttt{I} & ::= & \texttt{Reg := Val} \\
                 & | & \texttt{Loc := Val} \\
                 & | & \texttt{Reg := Loc} \\
                 & | & \texttt{CJmp Reg $\mathbb{N} \backslash \{0\}$} \\
  \texttt{Reg}            & ::= & r \, | \, r' \, | ... \\ 
  \texttt{Loc}            & ::= & x \, | \,y \,| ... \\ 
\end{array}
\] where
\begin{itemize}
  \item \texttt{P} -- program is a sequence of single thread progams
  \item \texttt{ST} -- single thread program is a sequence of instructions
  \item \texttt{I} -- instruction can be either write to some register of some value or location value, either write to some location or conditional jump
\end{itemize}
Here \texttt{Reg} are registers (denoted as $r,r',...$) ~--- variables that should be different for different single thread progams. In contrast \texttt{Loc} are local variables (denoted as $x,y,...$) ~--- this is our shared (between different threads) memory. Also \texttt{Val} is the type of values (denoted as $i,j,...$) in our program, for simplicity let it be $\mathbb{N}$ or $\mathbb{Z}$. \\
We are interested at looking after shared memory, so in our event structures we have events that represent actions only at local variables. \\
Next we define semantics of single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define thread state:
\begin{definition}[Thread State]
  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \regmap \rangle,$$ where
  \begin{itemize}
    \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
    \item $\regmap$ ~--- map from registers to values
  \end{itemize}
\end{definition}
If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$ (we start numeration from 1).  \\
Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where
\begin{itemize}
  \item $st$ ~--- thread state
  \item $p$ ~--- our single-thead programm
  \item $l$ ~--- label that denotes action on shared memory (can be with hole)
  \item $cont\_st$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
\end{itemize}
If we have some thread state $st$, we can identify $\lambda\, x,\, st$ (ie. constant continuation) with $st$ itself. Also by $f[r \leftarrow i]$ we will denote function that sends $r$ to $i$ and $r'$ to $f(r')$, for all $r' \neq r$
